---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

I unzip the file into the working directory. I read the data in and assign it as "act".

```{r gettingthedata, echo=TRUE}
 unzip("activity.zip")
 act <- read.csv("activity.csv")
```

I summarize the data to check for NA values and clean them.

```{r summary, echo=TRUE}
summary(act)
act <- na.omit(act)
summary(act)
```

## What is mean total number of steps taken per day?
There are 53 unique dates on which data was recorded.

```{r, echo=TRUE}
length(unique(act$date))
```

I load the package dplyr in order to process the data.Then I group the data under dates and calculate the total number of steps taken per day, storing the results in "daily_steps".

```{r steps, echo=TRUE}
library(dplyr)
daily_steps <- summarise(group_by(act, date), total = sum(steps))
```

I create a histogram of the total number of steps taken each day. 

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
