---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


## Loading and preprocessing the data

I unzip the file into the working directory. I read the data in and assign it as "act".

```{r gettingthedata, echo=TRUE}
 unzip("activity.zip")
 act <- read.csv("activity.csv")
```

I summarize the data.

```{r summary, echo=TRUE}
summary(act)
```

## What is mean total number of steps taken per day?
There are 53 unique dates on which data was recorded.

```{r, echo=TRUE}
length(unique(act$date))
```

I load the package dplyr in order to process the data.Then I group the data under dates and calculate the total number of steps taken per day, storing the results in "daily_steps".

```{r steps, echo=TRUE}
library(dplyr)
daily_steps <- summarise(group_by(act, date), total = sum(steps))
```

I create a histogram of the total number of steps taken each day.

```{r histogram, echo=TRUE}
hist(daily_steps$total, main="Steps per Day", xlab="Number of steps", col = "blue")
```

I calculate the mean and median of the total number of steps taken per day.

```{r distribution, echo=TRUE}
mean(daily_steps$total, na.rm = TRUE)
median(daily_steps$total, na.rm = TRUE)
```

## What is the average daily activity pattern?

I process the data and take the average of steps taken in each 5-minute interval across all days.
```{r intervals, echo=TRUE}
by_intervals <- aggregate(steps ~ interval, act, mean)
```

I construct a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).

```{r timeseries, echo=TRUE}

plot(by_intervals$interval, by_intervals$steps, type='l', main="Daily Average of Activity by 5-Min Intervals", xlab="5-Min Interval", ylab="Average number of steps", col="red")
```

I find out the 5 minute interval with the highest daily average.

```{r maxinterval, echo=TRUE}
which.max(by_intervals$steps)
by_intervals[104, ]
```

## Imputing missing values

I calculate the number of missing values.

```{r missing, echo=TRUE}
sum(is.na(act))
```

My strategy to impute the NA values is to replace them with each corresponding interval's mean. The average number of steps for each interval is already stored in "by_intervals". I make R replace each NA value in "act" with the corresponding mean from the "by_intervals" data frame.

```{r impute, echo=TRUE}
for (i in 1:nrow(act)){
  if (is.na(act$steps[i])){
    interval_no <- act$interval[i]
    row_no <- which(by_intervals$interval == interval_no)
    steps_ave <- by_intervals$steps[row_no]
    act$steps[i] <- steps_ave
  }
}
```

I create a histogram of daily steps using the imputed data.

```{r imputedhist, echo=TRUE}
imputed <- aggregate(act$steps ~ act$date, act, sum)
hist(imputed$`act$steps`, col="blue", main = "Histogram with NA Values Replaced", xlab="Daily Steps")
```
Calculating the mean and median of the new data, I observe that the mean did not change from what it was before NA values were imputed and the median is now equal to that value.
```{r imputeddist, echo=TRUE}
mean(imputed$`act$steps`)
median(imputed$`act$steps`)
```

## Are there differences in activity patterns between weekdays and weekends?

I extract the day information using the "weekdays" function and store it in a new object named "days".

```{r days, echo=TRUE}
days <- weekdays(as.Date(act$date))
```

I create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day. Using the "days" vector to seperate the between the two levels.

```{r wd_or_we, echo=TRUE}
wd_or_we <- vector()
for (i in 1:nrow(act)) {
  if (days[i] == "Cumartesi") {
    wd_or_we[i] <- "Weekend"
  } else if (days[i] == "Pazar") {
    wd_or_we[i] <- "Weekend"
  } else {
    wd_or_we[i] <- "Weekday"
  }
}

act_by_day <- aggregate(steps ~ interval + wd_or_we, data = act, FUN = mean)

```


I load the lattice package. I make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r we_or_wd_plot, echo=TRUE}
library(lattice)
xyplot(steps ~ interval | wd_or_we, act_by_day, type = "l", layout = c(1, 2))
```
